---
title: "Assignment 4"
author: "Lory Salazar"
date: "11/17/2020"
output: html_document
---

```{r setup, echo=TRUE, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(lubridate)
```

## Introduction
intro here

## Data and analyses
data here

## Exploratory findings

Read in the dataset first, and then begin data wrangling.
```{r}
bonanza_hares <- read.csv(here("data","bonanza_hares.csv"))
```

### Annual juvenile hare trap counts plot

First, I add a column to include the year. After, I use the count and filter functions to create a subset of the data that has counts for the juvenile hares. Lastly, I created a bar graph to plot the number of juvenile hares per year.
```{r}
bonanza_hares_dates <- bonanza_hares %>% 
  mutate(date_updated = mdy(date)) %>% 
  mutate(year = year(date_updated))

age_count <- bonanza_hares_dates %>% 
  count(age, year) %>% 
  filter(age == "j")

ggplot(data = age_count, aes(x = year, y = n)) +
  geom_col() +
  labs(x = "Time (years)",
       y = "Number of juvenile hare trappings",
       title = "Total number of juvenile hare trappings by year (Bonanza Creek)") +
  scale_y_continuous(breaks = c(20,40,60,80,100,120,140)) +
  theme_minimal() 
```

Explain/describe major takeaways here:


### Juvenile hare weights plot


